---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`
```{r}
library(tidyverse)
library(readr)
library(dplyr)
```

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 
```{r}
returns <- av |> 
  select(Name.Alias, starts_with("Return")) |>
  pivot_longer(
    Return1:Return5,
    names_to = "Time",
    values_to = "Returned"
  ) |>
  mutate(
    Time = parse_number(Time)
  )



head(returns)

deaths <- av |> select(Name.Alias,
    starts_with("Death")) |>
  pivot_longer(
    Death1:Death5,
    names_to = "Time",
    values_to = "Died"
  ) |>
  mutate(
    Time = parse_number(Time)
  )

head(deaths)


totaldeaths = nrow(filter(deaths, Died=='YES'))
#totaldeaths

totalavengers = length(unique(deaths$Name.Alias))
#totalavengers

avg = totaldeaths/totalavengers
avg

```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

Allison: "I counted 89 total deaths"

Ivy: "My analysis found that 69 had died at least one time after they joined the team."

Norah: "on 57 occasions the individual made a comeback"

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

Allison: 

Ivy: 
```{r}
died_once <- deaths %>%
  mutate(died = Died == "YES") %>% 
  group_by(Name.Alias) %>%
  summarise(n_deaths = sum(died, na.rm = TRUE), .groups = "drop") %>%
  summarise(n = sum(n_deaths >= 1))



died_once
```


Norah: 
```{r}

head(returns)

comebacks <- count(returns, Returned=='YES')
comebacks
newcb <- filter(comebacks, comebacks$`Returned == "YES"`==TRUE)
newcb

```

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

Allison:

Ivy: Based on my analysis of the dataset, **64 Avengers** have died at least one time after joining the team. This number is slightly lower than the FiveThirtyEight claim of 69 deaths. The difference may be due to dataset updates or variations in how deaths were counted.

Norah: Based on the code that I used to analyze the returns dataset, there were in fact 57 times that an individual died and returned. This exactly matches the number of times that the article claims people returned from the dead.
